Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-08-19 paper/main.tex
rewrote top def

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total, two heads] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A continuous surjective map \textbf{$\pi$ is bundle projection} map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$. All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$. 
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}. For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on $\openset$}, is locally trivial. The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
\end{enumerate}
\end{definition}
We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber} and a corresponding index, as discussed in \autoref{sec:atct:fb:base}. For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.

@@

\subsubsection{Topological Structure: Base Space \dbase}
\label{sec:atct:fb:base}
We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}. The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval. This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}. An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$. Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.

\begin{definition}
A \textbf{topology} $\mathcal{T}_{\dbase}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
  \begin{itemize}
    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}_{\dbase}$
    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the intersection $\openset_i \cap \openset_j$  is in $ \mathcal{T}_{\dbase}$
    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the union $\openset_i \cup \openset_j$ is in $ \mathcal{T}_{\dbase}$
  \end{itemize}
  such that the the topology provides an axiomatic way to construct the topological space $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$\cite{bradleyTopologyCategoricalApproach2020}.
\end{definition}
 A property of the base space \dbase\ of a fiber bundle is that it is a quotient topology, which  means that it contains the largest possible collection of distinct opensets, meaning each openset is as small as possible, of $\dbase$ that still preserve the underlying continuity of \dbase \cite{QuotientSpaceTopology2020}. For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}. We express the allowable ways to subset and compose points such that the connectivity is preserved as the topology $\mathcal{T}_{\dbase}$ on the data continuity.

@@

When the total space is non-trivial, we can use the fiber bundle property of local-triviality to generate type signatures. In a non-trivial bundle, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$. Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood $\dbasepoint \in \openset \in \dbase$. This is because all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal. This means almost all data can be expressed as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ which translate to a set of signatures 
\begin{equation}
  \begin{split}
\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}} \\
 s.\;t.\;\textcolor{section}{\texttt{data-subset}} \in \texttt{dataset}\}  
  \end{split}
\end{equation}
that can be glued together on overlaps to generate the dataset. This gluing can be implemented as part of the data container, which we discuss in \autoref{sec:atct:sheaves}.

>>>>

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total, two heads] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A continuous surjective map \textbf{$\pi$ is bundle projection} map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$. All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$. 
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}. For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on the points in $\openset$}, is locally trivial. The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
\end{enumerate}
\end{definition}
We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}. For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.

@@

\subsubsection{Topological Structure: Base Space \dbase}
\label{sec:atct:fb:base}
We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}. A topology $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$ is an axiomatic way of defining a topological structure\cite{bradleyTopologyCategoricalApproach2020}  as a collection of subsets, called open sets \footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces. For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball. \cite{weissteinOpenSet}}, of that structure.

\begin{definition}
A \textbf{topology} $\mathcal{T}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
  \begin{itemize}
    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$
    \item the union of open sets in $\mathcal{T}$ is in also in $\mathcal{T}$
    \item any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$
  \end{itemize}
\end{definition}
The base space of a fiber bundle is a quotient topology\cite{QuotientSpaceTopology2020}, meaning that it divides the topological space into the smallest open sets such that $\pi$ remains a continuous function.  For our purposes, $\dbase$ encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}. We define $\dbase$ as a topology $\mathcal{T}_{\dbase}$ because a topology codifies that the underlying topological structure of the data can be subsetted and how these subsets can be composed in a structure preserving manner.

@@

When the total space is non-trivial, we can use the fiber bundle property of local-triviality to generate type signatures. In a non-trivial bundle, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$. Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood\footnote{An open set that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.} $\dbasepoint \in \openset \in \dbase$. This is because all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal. This means almost all data can be expressed as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ which translate to a set of signatures 
\begin{equation}
  \begin{split}
\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}} \\
 s.\;t.\;\textcolor{section}{\texttt{data-subset}} \in \texttt{dataset}\}  
  \end{split}
\end{equation}
that can be glued together on overlaps to generate the dataset. This gluing can be implemented as part of the data container, which we discuss in \autoref{sec:atct:sheaves}.
